---
author: "Santiago Patiño-Giraldo"
date: "21/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This work will present the analysis for an anonymous subject about his/her daily
movement activities.

## Preparing the data

A couple of packages for the work are loaded at first
```{r results='hide'}
library(lubridate)
library(dplyr)
library(xtable)
```

First, we need to unzip the dataset -if wasn't made previously-

```{r}
if(!file.exists("activity.csv")){
        unzip(zipfile = "activity.zip")
}
```

Then, the data is loaded

```{r}
activity <- read.csv("activity.csv", header=TRUE)
```

It's important to check how the data is looking reviewing the head of dataset and
the class of the columns
```{r}
head(activity)
str(activity)
```

I'm not happy with the class of data, so, I will change them

```{r}
activity$date <- ymd(activity$date)
activity$interval <- as.factor(activity$interval)
```

##What is mean total number of steps taken per day?

To solve this question I group the data by date and then summarize the mean
```{r}
meansteps <- group_by(activity[complete.cases(activity),],date) %>% 
                summarize(daily.steps=sum(steps),
                mean_day=mean(steps,na.rm = TRUE),
                median.day=median(steps))
htmltable <- xtable(meansteps)
```

And the histogram of the total steps

```{r fig.align='center'}
hist(meansteps$daily.steps, main="Histogram of Mean Steps", ylim=c(0,30), xlab = "Mean steps by day", col=2)
```
###Summary of data
```{r results='asis'}
print(htmltable, type="html")
```
